Rapid technological progress in white light-emitting diodes \\ and its sources in innovation and technology spillovers

### 2023-08-12 2_SSL_NE/article/article_main.tex
added edits

>>>>

The evolution of LED device architecture and performance, as well as the progress in understanding the underlying physical phenomena, are relatively well covered in scholarly literature and patents (see Supplementary Note 1 for a brief literature review). However, information provided in such sources is insufficient for our goals on at least three accounts: First, existing work focuses only on selected performance parameters or overall device efficiency, rather than on providing a comprehensive coverage of the whole device sub-efficiencies for a particular LED product or design. Scientific publications also do not always disclose the underlying device architecture or the features responsible for the gains in performance. Second, not all relevant innovations are patented \cite{Pakes_1980,Fontana_2013}. In the case of LED patents in particular, our interviews with industry experts suggest that the propensity to patent is the highest for knowledge related to macroscopic device architecture and chemical composition of phosphors, and the lowest for knowledge related to manufacturing process improvements and microscopic chip architecture that is difficult to reconstruct by reverse engineering. This means that relying only on patent literature would bias results by unduly emphasizing some focus areas and de-emphasizing others. Third, scientific publications and patents typically focus on experimental devices, rather than commercial products. While new LED features, designs and manufacturing methods reported in these sources can potentially result in significant performance gains or cost reductions, it is difficult to ascertain if these improvements have since been adopted in industry. Furthermore, information on LED manufacturing cost and the effect of process improvements on the total cost is highly proprietary. Estimates are occasionally reported in the scientific literature and company publications, but these often do not disclose which parts of the manufacturing process are responsible for the largest contribution to the overall cost, or which improvements led to cost reductions.

To overcome the limitations of existing literature and patent analysis, in this study we rely on a multi-method approach to data collection and analysis, augmenting information obtained from a systematic review of the primary scientific literature, device datasheets, relevant patents, and industry publications with information gained from semi-structured interviews with experts from academia and industry, bottom-up manufacturing cost modelling, and own calculations of LED device sub-efficiencies. We then synthesize information obtained with multiple methods to track the historical progress in white LED technology over time across the three groups of metrics (see Supplementary Note 2) and identify its sources in innovation and technology spillovers. We briefly describe our approach to each method below and provide further methodological details in Supplementary Note 3.

\subsection{Systematic Literature Review}

We collected data on LED performance and characteristics in a systematic literature review that included scientific publications, patents, conference proceedings from the largest semiconductor and optoelectronics conferences, industry periodicals and roadmaps, as well as company presentations and reports. This review was structured around the three main goals: 1) tracking the evolution of LED technology over time as indicated by three groups of progress metrics; 2) identifying individual innovations that contributed to this evolution, and quantifying their impact on device performance and manufacturing cost; and 3) determining whether these innovations had originated within the LED technology domain, or in a field of science or technology outside of SSL, making them involved in technology spillovers.

Relevant sources for the review were found in an iterative search process that involved two components. The first was the search in specialized patent and publication databases as well as company websites. The second component was the analysis of backward citations in the identified sources, starting from the reviews mentioned in Supplementary Note 1 and then iteratively repeating it for all newly identified sources, until no further relevant and significant new sources were found. We also relied on backward citations in these sources for the identification of technology spillovers, considering cited documents as indicators of knowledge origins of an innovation and analyzing whether those documents belonged to the LED technology domain or not.

\subsection{Semi-Structured Interviews}

To supplement our data collection efforts, verify our findings and identify additional spillovers, we conducted a series of elite semi-structured interviews \cite{tansey2009process} with thirteen eminent experts from academia, industry, and public research sector. Experts were initially selected based on their engagement in different sub-fields of LED research and manufacturing, based on the literature review, then the list was expanded by a ‘snowballing’ tactic based on recommendations from already-interviewed experts. All interviews were conducted between November 2019 and April 2022 by means of video conferencing and lasted for about one hour. An anonymized summary of the background of interviewed experts is provided in Supplementary Table 1. Notably, all our interviewed experts came from Europe or USA, with none representing Asia, which potentially may have biased our findings particularly for the earliest and latest periods of white LED history dominated by LED manufacturers in Japan and China, correspondingly. However, this was not an intentional bias, as none of the identified experts in Asia responded to our interview requests.

In terms of the interview content, the primary, structured part of the interviews explored which innovations were deemed most relevant to the evolution of device performance, consumer experience and manufacturing cost of LED packages. Thereafter, interviewees were asked to consider the extent to which those innovations may have originated outside of their respective field of expertise and the LED industry more broadly—i.e., which of the innovations potentially involved technology spillovers. The remainder of the interview was focused on learning about particular aspects of the manufacturing processes relevant to cost and performance modelling, the current state of industry, and the circumstances surrounding the innovations and spillovers identified in the first part of the interview. Specific quantitative data was also provided by experts, helping fine-tune the parameters of the manufacturing cost model and verify device performance data.

\subsection{Performance Metrics Calculations}

The contribution of individual technology innovations and spillovers to the progress in overall device efficiency over time is estimated by index decomposition analysis. Mathematically, this involves breaking down a chosen performance indicator into its constituent components, each representing a specific factor that contributes to the change in the indicator \cite{Ang1997}. Specifically, we use the additive logarithmic mean Divisia index method I (LMDI-I), also known as the Additive Sato-Vartia indicator \cite{deBoer2019}. It was developed by Boyd in 1987 \cite{Boyd1987} on the basis of Divisia Index, a method in statistical economics \cite{Diewert1988}, and subsequently refined.

According to this method, for an overall device efficiency function $F$ that is the product of variables $a, b$ that represent sub-efficiencies, the contribution of the change in a single sub-efficiency variable $a$ between times $t=0$ and $t=T$ can be estimated as \cite{Ang2019}
%
\begin{align}
    \Delta a &= \frac{a_{t=T} - a_{t=0}}{\ln(a_{t=T}) - \ln(a_{t=0})} \times \ln \big ( \frac{a_{t=T}}{a_{t=0}} \big ) \\
    & \stackrel{a_{t=0} \neq a_{t=T}}{=} L(F_{t=T}, F_{t=0}) \times \ln \big ( \frac{a_{t=T}}{a_{t=0}} \big )
\end{align}
%
where $L(F_{t=T}, F_{t=0})$ is the logarithmic mean of $F$ values at times $t=0$ and $t=T$. These terms contain no residuals, therefore it can be shown that the overall improvement in the device efficiency due to improvements in individual sub-efficiencies is equal to the sum of these improvements in individual sub-efficiencies: 
%
\begin{equation}
    \Delta a + \Delta b  = \Delta F
\end{equation}
%
To document historical improvements in LED device performance accurately, we need data on all sub-efficiencies for the selected device architectures and periods covered. However, the scope of data reporting in the literature is typically limited to selected metrics of interest, rather than the full ensemble of sub-efficiencies that determine the overall device performance. Where gaps in relevant data existed, we filled them with our own performance calculations for individual sub-efficiencies where possible. Specifically, sub-efficiencies related to the emission spectrum of phosphor down-converted LED devices (i.e., spectral efficiency and light conversion efficiency)  were computed from the spectral data, often reported in LED device specifications, using the \texttt{colour-science} package for Python \cite{colour-science_software}. We also used the same approach to calculate CRI and luminous efficacy of radiation of LED devices (see Supplementary Note 4 and Supplementary Figure 10 for the associated spectral data and calculation results). This approach allowed us to quantify the improvements related to consumer experience and phosphor development in LEDs.

\subsection{Manufacturing Cost Model}

The structure of our bottom-up white LED manufacturing cost model with process-step resolution is generally based on the 2012 \textit{LEDCOM} cost model \cite{ledcomv2}, but we expand it significantly both in scope and in its ability to capture historical trends. The model captures three historical time periods corresponding to different “eras” in LED manufacturing: the early period of the first high-power white LEDs around 2003, the period of accelerating consumer adoption of LED lighting around 2012, and the most recent period around 2020, the year of our main data collection efforts. For each of these three years, the most prevalent manufacturing equipment was identified through industry periodicals, archived website data from the Internet Archive , and expert interviews. Because the architecture of LED chips has changed significantly since the introduction of the first commercial white LED devices in 1996, three different chip architectures are considered in the model: classical chips, flip chips, and chip-scale package flip chips (see Figure 3 and Supplementary Figures 2-6 for the details of each architecture). The details of the manufacturing process for each architecture were collected from the scientific literature, textbooks  and relevant patents. In addition, two LED life cycle analyses (Scholand and Dillon 2012a)(Casamayor, Su, and Ren 2018) were used to validate the model structure and extract some of the necessary quantitative model inputs. These studies captured a large number of LED manufacturing process steps and included the details on the use of metals, chemicals and electricity for each manufacturing step.

The cost model we developed is based on a cumulative approach to yielded cost\cite{becker2001use}. In this approach, the yielded cost $C_{Y_i}$ of process step $1$ is defined as the ratio between the total cost of step 1 $C_1$ and the yield of step 1 $Y_1$. Thus, for each consecutive step, starting from $i=1$, we get
%
\begin{equation}
    C_{Y_1} = \frac{C_1}{Y_1}, \ C_{Y_2} = C_{Y_{2 \rightarrow 3}} - C_{Y_1} = \frac{C_1(1-Y_2)+C_2}{Y_1Y_2}, \ C_{Y_3}=\dots
\end{equation}
%
Notably, yielded cost per step is dependent on the step order and blind to downstream information \cite{becker2001use}. The yielded cost metric is also cumulative by definition, thus the total cumulative yielded cost is calculated as:
%
\begin{equation}
    \sum_i C_{Y_i} = \frac{\sum_i C_i}{\prod_i Y_i}
\end{equation}
%
The overall outcome of the cost model is thus the cumulative yielded  manufacturing cost per LED package for each of the three years considered. In our model, this metric includes all costs associated with producing the chip, including running costs of the factory. Costs associated with research and development, administrative overhead of the manufacturer or other investment costs are not considered. We also note that the purpose of our cost model is not to give specific estimates of the LED manufacturing cost for a factory of any size, specific geographic location or total annual manufacturing volume. It instead assumes a hypothetical factory with an assumed location in the United States and associated overhead costs related to the operation of the factory. It also assumes the use of the most up-to-date equipment for that year. Even with these simplifying assumptions, the model reasonably identifies the impact that changes in single process steps can have on the total LED manufacturing cost.

Another important limitation of our cost modelling efforts is that, even though the model captures three different chip architectures in its structure, in the present study we are able to collect, estimate and present the full set of quantitative inputs and outputs only for the classical chip architecture of low- to mid-power devices. Populating the model with data for the remaining two architectures requires access to proprietary industry information, which we have not been able to get thus far.

Further details on our manufacturing cost model, including its structure and equations, manufacturing process flows for the chip architectures under consideration, input data, detailed calculations for the yielded costs, as well as the model’s limitations, are provided in Supplementary Note 3.

### 2023-08-12 2_SSL_NE/article/article_supplementary.tex
added edits

Notably, \cref{fgr:subeff} does not include sub-efficiencies related to spectral performance of LEDs: spectral efficiency $\eta_S$ and conversion efficiencies of different phosphors $η_{C_R}$ and $\eta_{C_{Y/G}}$. Typically, there is no direct empirical data available for these metrics in the literature, which made their inclusion in the empirical \cref{fgr:subeff} not relevant. Instead, we calculate the values of these metrics used in Figures 4 and 5 in the main text from spectral and luminous efficacy radiation data presented in \cref{fig:phosphor_spectrum}.

>>>>

Notably, \cref{fgr:subeff} does not include sub-efficiencies related to spectral performance of LEDs: spectral efficiency $\eta_S$ and conversion efficiencies of different phosphors $\eta_{C_R}$ and $\eta_{C_{Y/G}}$. Typically, there is no direct empirical data available for these metrics in the literature, which made their inclusion in the empirical \cref{fgr:subeff} not relevant. Instead, we calculate the values of these metrics used in Figures 4 and 5 in the main text from spectral and luminous efficacy radiation data presented in \cref{fig:phosphor_spectrum}.
